---
title: Applied Statistical Methods - `r tools::toTitleCase(params$doctype)` 3
author: Peter von Rohr
date: 2021-03-15
output:
  bookdown::pdf_document2:
    toc: false
    number_sections: false
    keep_tex: false
header-includes:
 \usepackage{longtable}
 \usepackage{float}
params:
  doctype:
    label: 'Document Type'
    value: 'exercise'
    choices: ['exercise', 'solution']
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = 'asis', fig.pos = 'H')
knitr::knit_hooks$set(hook_convert_odg = rmdhelp::hook_convert_odg)
```


## Problem 1: Fixed Linear Effects Model {-}
We want to analyse a dataset with genetic information using a fixed linear effects model. The dataset is taken from the course notes and is shown in Table \@ref(tab:datatable). 

We assume that the SNP loci have a purely additive effect on the trait. That means for a SNP locus $L$ the absolute value of the genotypic value of the homozygous genotypes ($L_1L_1$ and $L_2L_2$) is taken to be $a_L$ and the genotypic value of the heterozygous genotype ($L_1L_2$) is taken to be $0$. The fixed linear effects model contains the observation in Table \@ref(tab:datatable) as the response variable, an intercept and the genotypic values of the the genotypes at the two SNP Loci $G$ and $H$ as predictor variables.

For the observation $y_i$ of animal $i$, we can specify the model as

$$y_i = \beta_0 + W_i \cdot a + \epsilon_i$$

where $\beta_0$ is the intercept, $a$ is the vector of additive SNP-effects, $W_i$ is a row vector denoting the SNP-Genotypes and $\epsilon_i$ is the random error term.

```{r dataurldeclare, echo=FALSE}
### # specify path to data file depending on online status
s_data_file <- "https://charlotte-ngs.github.io/gelasmss2021/data/ex03p01_data.csv"
```

The data can be read from `r s_data_file`. The address from where the data can be downloaded is assigned to a variable. 

```{r dataurldeclare, echo=TRUE}
```


It can be read using the following statement

```{r datainput, message=FALSE}
### # read the data into a tibble
tbl_geno_data <- readr::read_csv(file = s_data_file)
```

```{r datatable, echo=FALSE, results='asis'}
knitr::kable(tbl_geno_data,
             booktabs = TRUE, 
             longtable = TRUE, 
             caption = "Genotypic Data Used for Fitting a Fixed Linear Effect Model")
```

### Your Tasks {-}
* Specify the fixed linear effects model in matrix-vector notation by putting the information from the dataset into the model. Use the same parametrization as shown in the course notes where the intercept $\beta_0$ and the vector $a$ are combined into a single parameter vector $b$. The design matrix that links elements in $b$ to observations $y$ is then called $X$. 
* Use the function `Matrix::rankMatrix()` from the `Matrix` package on the matrix $X$ to find out the rank of the design matrix.
* Depending on the rank of $X$ compute an estimate for $b$, if the rank of the matrix is equal to the number of columns of matrix $X$, then the same forumla as was used in the regression model can be used
* Verify your results using the `lm()` function


### Hints {-}
* Read the data using the function `readr::read_csv()`


```{r p01-comment-start, echo=FALSE, results='asis'}
if (params$doctype != 'solution')
  cat('<!-- \n')
```

### Solution {-}

__Model Specification__: The fixed linear effect model in matrix-vector notation is given by 

$$y = X\cdot b + e$$
where the vector $b$ contains all unknown parameters which means, $b = \left[\begin{array}{c}\beta_0 \\ a \end{array} \right]$ and $X$ is the design matrix consisting of a column of all ones and with one column for each SNP locus. For our example we have 

$$b = \left[\begin{array}{c}\beta_0 \\ a_G \\ a_H \end{array} \right]$$

The matrix $X$ comes from the data. The first columns of $X$ is all ones and the second and the third column are codes corresponding to $-1$, $0$ or $1$, depending on the genotypes of the animals at the SNP loci $G$ and $H$. 

```{r}
n_nr_animal <-  nrow(tbl_geno_data)
mat_X <- matrix(c(rep(1, n_nr_animal), tbl_geno_data$`SNP G`, tbl_geno_data$`SNP H`), ncol = 3)
```

The matrix $X$ corresponds then to

```{r showmatx, echo=FALSE, results='asis'}
cat(paste0(rmdhelp::bmatrix(pmat = mat_X, ps_name = 'X', ps_env = '$$'), collapse = '\n'), '\n')
```

The vector $y$ corresponds to 

```{r showvecy, echo=FALSE, results='asis'}
vec_y <- tbl_geno_data$Observation
cat(paste0(rmdhelp::bcolumn_vector(pvec = vec_y, ps_name = 'y', ps_env = '$$'), collapse = '\n'), '\n')
```


__Rank of Matrix $X$__: The rank of the matrix $X$ is obtained by

```{r rankmatrixx, results='markup'}
Matrix::rankMatrix(mat_X)
```

__Solution for $\hat{b}$__: Because matrix $X$ has full column rank which means the rank of the matrix is the same as the number of columns, the solution for $\hat{b}$ can be computed the same way as for the regression model. Hence

$$\hat{b} = (X^TX)^{-1}X^Ty$$

In R this corresponds to 

```{r solhatb, results='markup'}
(vec_hatb <- crossprod(solve(crossprod(mat_X)), crossprod(mat_X, vec_y)))
```

__Verify Results in R__:

```{r lmsnpflem, results='markup'}
lm_snpflem <- lm(Observation ~ `SNP G` + `SNP H`, data = tbl_geno_data)
summary(lm_snpflem)
```

```{r p01-comment-end, echo=FALSE, results='asis'}
if (params$doctype != 'solution')
  cat('--> \n')
```


## Problem 2: Genomic Relationship Matrix {-}
```{r declaredataex03p02, echo=FALSE}
s_data_ex03p02 <- "https://charlotte-ngs.github.io/gelasmss2021/data/ex03p02_data.csv"
```

From the given dataset that can be obtained from 

`r s_data_ex03p02`, 

compute the genomic relationship matrix $G$. The dataset is organised such that animals are in rows and SNPs are in columns.


### Hints {-}
* Read the data using the function `readr::read_csv()`
* Convert the input data with the function `as.matrix()` to a matrix
* Use the function `apply(mat, 2, mean)` to compute the columnwise mean of matrix `mat`
* Use the `matrix` function to construct the matrix $P$ from the vector of SNP allele frequencies

```{r p02-comment-start, echo=FALSE, results='asis'}
if (params$doctype != 'solution')
  cat('<!-- \n')
```

### Solution {-}
Start by assigning the address of the input file to a variable.

```{r declaredataex03p02, echo=TRUE}
```

Now the data can be read from the file. 

```{r readdataex03p02, message=FALSE}
tbl_grm_data <- readr::read_csv(file = s_data_ex03p02)
n_nr_animal <- nrow(tbl_grm_data)
n_nr_snp <- ncol(tbl_grm_data)
```

The frequency of the positive alleles for all SNP positions is computed as 

```{r}
mat_W <- as.matrix(tbl_grm_data)
vec_allele_freq <- apply(mat_W+1, 2, mean)/2
```

The sum of $p_jq_j$ is computed as

```{r}
sumpq <- sum(vec_allele_freq * (1-vec_allele_freq))
```

The matrix $U$ is computed from the matrix $W$ and the matrix $P$

```{r}
mat_P <- matrix(2*vec_allele_freq-1, nrow = n_nr_animal, ncol = n_nr_snp, byrow = TRUE)
mat_U <- mat_W - mat_P
```

The genomic relationship matrix is obtained by

```{r}
mat_grm <- tcrossprod(mat_U) / (2*sumpq)
```

```{r p02-comment-end, echo=FALSE, results='asis'}
if (params$doctype != 'solution')
  cat('--> \n')
```


## Additional Problem {-}
Write a function in R that accepts a matrix of genotypes and that computes the genomic relationship matrix. Verify your results from Problem 2.

```{r addp01-comment-start, echo=FALSE, results='asis'}
if (params$doctype != 'solution')
  cat('<!-- \n')
```

### Solution {-}
The following function computes the genomic relationship matrix

```{r}
#' Compute genomic relationship matrix based on data matrix
computeMatGrm <- function(pmatData) {
  matData <- pmatData
  # check the coding, if matData is -1, 0, 1 coded, then add 1 to get to 0, 1, 2 coding
  if (min(matData) < 0) matData <- matData + 1
  # Allele frequencies, column vector of P and sum of frequency products
  freq <- apply(matData, 2, mean) / 2
  P <- 2 * (freq - 0.5)
  sumpq <- sum(freq*(1-freq))
  # Changing the coding from (0,1,2) to (-1,0,1) and subtract matrix P
  Z <- matData - 1 - matrix(P, nrow = nrow(matData), 
                             ncol = ncol(matData), 
                             byrow = TRUE)
  # Z%*%Zt is replaced by tcrossprod(Z)
  return(tcrossprod(Z)/(2*sumpq))
}
```

The function is tested with

```{r, results='markup'}
mat_grm_func <- computeMatGrm(pmatData = mat_W)
all.equal(mat_grm, mat_grm_func)
```


```{r addp01-comment-end, echo=FALSE, results='asis'}
if (params$doctype != 'solution')
  cat('--> \n')
```

