---
title: Applied Statistical Methods --  Notebook 2
author: Peter von Rohr
date: 2021-03-08
output: html_notebook
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = FALSE, results = 'asis')
if (!is.element('rmdhelp', installed.packages()))
  remotes::install_github(repo = 'charlotte-ngs/rmdhelp')
knitr::knit_hooks$set(hook_convert_odg = rmdhelp::hook_convert_odg)
```


## Problem 1: Multiple Linear Regression Model {-}
During the lecture the regression model was explained using the dataset on body weight and breast circumference. To improve the regression model for body weight, two additional conformation traits `BCS` and `HEI` are also considered. The new dataset is given in Table \@ref(tab:dataregression). 

```{r dataregression, echo=FALSE, results='asis', message=FALSE}
### # define constant paths
s_data_root_online <- "https://charlotte-ngs.github.io/gelasmss2021/data"
s_data_dir <- "asm_w03_ex02_bw_mult_reg.csv"
s_data_root <- s_data_root_online
s_data_path <- file.path(s_data_root, s_data_dir)
### # read data from s_data_path
tbl_reg <- readr::read_csv(file = s_data_path)
knitr::kable(tbl_reg,
             booktabs = TRUE,
             longtable = TRUE,
             caption = "Dataset for Multiple Linear Regression of Body Weight on Breast Circumference and two Conformation Traits for ten Animals")
```

The same dataset is also available from the website at `r paste(s_data_root_online, s_data_dir, sep = "/")`. 



### Your Task {-}
* Setup the linear regression model with an intercept for the data given in Table \@ref(tab:dataregression)
* Compute the solution for the unknown parameter $b$
* Verify the result with the output from the function `lm()` in R


### Your Solution
To setup the linear regression model, we have to determine two components, namely the matrix $X$ and the vector $y$

```{r}
# specify the matrix X

# specify the vector y

```



## Problem 2: Prediction {-}
Given the measurement of the trait `Breast Circumference` for two additional animals. The measurements are shown in the following table

```{r tablenewanimal, echo=FALSE, results='asis'}
tbl_new_ani <- tibble::tibble(Animal = c("Animal 11", "Calf 12"),
                              `Breast Circumference` = c(181.2, 99.5))
knitr::kable(tbl_new_ani, 
             booktabs = TRUE, 
             longtable = TRUE, 
             caption = "Breast Circumference Measurements For Two Animals Used To Predict Body Weight")
```

We want to use the results of the simple linear regression of body weight (BW) on breast circumference (BC), as shown in Problem 2 of Exercise 1 to computed the predicted values for `Body Weight` for the two animals. The observed value for `Breast Circumference` of "Calf 12" is outside of the range of the values used in Problem 1. Predicting values of response variables based on predictors that are outside of the range of values used for the parameter estimation is called __extrapolation__. Based on the result of the predicted value of the trait `Body Weight` for "Calf 12" what can be said about the process of extrapolation? 


### Your Tasks {-}
* Compute the predicted value of `Body Weight` for "Animal 11" 
* Compute the predicted value of `Body Weight` for "Calf 12"
* Make a statement about the validity of the extrapolated value of `Body Weight` for "Calf 12"

 
