---
title: "Student Lists for GEL"
author: "Peter von Rohr"
date: "4/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Disclaimer
This document shows how the student lists for GEL are created. 

```{r}
vec_pkg <- c('openxlsx', 'dplyr', 'tidyr')
if (!all(is.element(vec_pkg, installed.packages())))
  install.packages(vec_pkg, dependencies = TRUE)

s_bl_path <- '~/Downloads/BelegungenLerneinheit_751621200L_Fruehjahrssemester_2021.xlsx'
wb <- openxlsx::read.xlsx(xlsxFile = s_bl_path)
```

First create a list of students with their names, firstnames and e-mail addresses

```{r}
require(dplyr)
tbl_student_info <- wb %>% 
    select(Familienname, Rufname, Nummer, `E-Mail`) %>%
    mutate(UDName = `E-Mail`) %>%
    tidyr::separate(UDName, c("Username", NA), sep = '@') %>%
    select(Username, Familienname, Rufname, Nummer, `E-Mail`)
tbl_student_info$Username <- tolower(tbl_student_info$Username)
tbl_student_info
```

The student information is written to a file. 

```{r}
s_student_dir <- file.path(here::here(), 'students')
if (!dir.exists(s_student_dir)) dir.create(s_student_dir)
readr::write_csv(tbl_student_info, path = file.path(s_student_dir, 'gel_students_gelasmss2021.txt'), col_names = FALSE)
```

A test-student is created manually. 

```{r}
tbl_test_student_info <- tibble::tibble(Username     = c('vrohrp'),
                                        Familienname = c('von Rohr'), 
                                        Rufname      = c('Peter'), 
                                        Nummer       = c('00-000-000') , 
                                        `E-Mail`     = c('peter.vonrohr@usys.ethz.ch'))
readr::write_csv(tbl_test_student_info, path = file.path(here::here(), 'students', 'test_gel_students_gelasmss2021.txt'), col_names = FALSE)
```

## Student Usernames
From the e-mail addresses, we extract usernames as follows

```{r}
(tbl_student_username <- tbl_student_info %>% select(Username))
```

Write the usernames to a file

```{r}
readr::write_csv(tbl_student_username, path = file.path(s_student_dir, 'gel_student_usernames_gelasmss2021.txt'), col_names = FALSE)
```

The username for the test-student is specified manually and is not extracted from the e-mail. 

```{r}
tbl_test_student_username <- tbl_test_student_info %>% select(Username)
readr::write_csv(tbl_test_student_username, path = file.path(here::here(), 'students', 'test_gel_student_usernames_gelasmss2021.txt'), col_names = FALSE)
```

## File Upload
Upload the created files to the server

```
$ cd /Users/peter/Data/Projects/GitHub/charlotte-ngs/gelasmss2021_gh-root/master/gelasmss2021
$ scp students/gel_student_usernames_gelasmss2021.txt quagadmin@2-htz.quagzws.com:/home/quagadmin/courses/gelasmss2021/students
$ scp students/test_gel_student_usernames_gelasmss2021.txt quagadmin@2-htz.quagzws.com:/home/quagadmin/courses/gelasmss2021/students
scp students/gel_students_gelasmss2021.txt  quagadmin@2-htz.quagzws.com:/home/quagadmin/courses/gelasmss2021/students
scp students/test_gel_students_gelasmss2021.txt  quagadmin@2-htz.quagzws.com:/home/quagadmin/courses/gelasmss2021/students

scp bash/docker_student_start.sh quagadmin@2-htz.quagzws.com:/home/quagadmin/courses/gelasmss2021/bash
scp bash/docker_student_stop.sh quagadmin@2-htz.quagzws.com:/home/quagadmin/courses/gelasmss2021/bash

```

## Starting Docker Containers

```
cd /home/quagadmin/courses/gelasmss2021
# test
./bash/docker_student_start.sh -f students/test_gel_students_gelasmss2021.txt -c /home/quagadmin/courses/gelasmss2021/home -p 10187 -i gelasmss2021 -w `pwd`/created
# real students
./bash/docker_student_start.sh -f students/gel_students_gelasmss2021.txt -c /home/quagadmin/courses/gelasmss2021/home -p 10096 -i gelasmss2021 -w `pwd`/created
```

## Creating Password Files

```
cd /home/quagadmin/courses/gelasmss2021
# test
cat students/test_gel_student_usernames_gelasmss2021.txt | while read u
do
  echo " * Extract pwd for $u ..."
  if [ ! -e "created/.${u}.pwd" ]
  then
    echo " ** Create pwd-file for $u: ..."
    grep 'Password' email_sent/${u}_email.txt  | cut -d ' ' -f 2 > created/.${u}.pwd
  else
    echo " ** Found pwd-file for $u: ..."
  fi
done
# real
cat students/gel_student_usernames_gelasmss2021.txt | while read u
do
  echo " * Extract pwd for $u ..."
  if [ ! -e "created/.${u}.pwd" ]
  then
    echo " ** Create pwd-file for $u: ..."
    grep 'Password' email_sent/${u}_email.txt   | cut -d ' ' -f 2 > created/.${u}.pwd
  else
    echo " ** Found pwd-file for $u: ..."
  fi
done

```


## Cloning

```
export GITURI=https://github.com/charlotte-ngs/gelasmss2021.git
export GELSRC=/home/quagadmin/courses/gelasmss2021
export GELPROJ=$(basename $GELSRC)
export GELHOME=${GELSRC}/home
cd $GELSRC
# test student
cat students/test_gel_student_usernames_gelasmss2021.txt | while read s
do
  cd $GELHOME/$s
  if [ -d "$GELPROJ" ]
  then
    echo " * $GELPROJ exists ..."
  else
    echo " * Cloning repo for $s ..."
    git clone $GITURI -b rexpf
  fi  
  sleep 2
done

# real
cat students/gel_student_usernames_gelasmss2021.txt | while read s
do
  cd $GELHOME/$s
  if [ -d "$GELPROJ" ]
  then
    echo " * $GELPROJ exists ..."
  else
    echo " * Cloning repo for $s ..."
    git clone $GITURI -b rexpf
  fi  
  sleep 2
done

# removed asm for new students

```

## Sending E-Mails
The following commands are used to send e-mails to students

```
cd /home/quagadmin/courses/gelasmss2021/bash
ls -1 | while read f
do
  echo " * Checking $f ..."
  if [ -e "../email_sent/$f" ]
  then
    echo " * $f exists ..."
    rm $f
  else
    echo " * $f new ..."
  fi
  sleep 2
done
```

