---
output:
  pdf_document:
    includes:
      in_header: tex/header.tex
    fig_caption: false
output_file: gelss2021_exam_sol_test.pdf
params:
  seed: 1234
  name: Name
  firstname: Firstname
  leginr: LegiNr
  alias: test
--- 
<!-- %\usepackage{fancyhdr} -->

\newcommand{\points}[1]
{\begin{flushright}\textbf{#1}\end{flushright}}

<!-- %\begin{document} -->
<!-- %\SweaveOpts{concordance=TRUE} -->
```{r ChunkOptions, echo=FALSE}
options(scipen = 9999)
# knitr::opts_chunk$set(echo = FALSE, results = 'hide')
#knitr::opts_chunk$set(concordance=TRUE)
knitr::knit_hooks$set(hook_convert_odg = rmdhelp::hook_convert_odg)
# write the parameters to file
b_params_to_file <- FALSE
# check whether seed is set and output it to a file
s_this_rmd_file = basename(ifelse(rstudioapi::isAvailable(), 
                         rstudioapi::getSourceEditorContext()$path, 
                         whereami::thisfile()))
if (is.null(params$seed)){
  stop(" ** Error parameter seed has not been set.")
} else {
  set.seed(params$seed)
  s_params_file <- paste0(format(Sys.time(), '%Y%m%d%H%M%S'), "_params_", s_this_rmd_file, ".txt", collapse = "")
  if (b_params_to_file) dput(params, file = s_params_file)
}
```

<!-- <<PointsQ1, echo=FALSE>>= -->
```{r PointsQ1, echo=FALSE}
# Assign Points for Q1
lPointsQ1 <- list(TaskA = 6,
                  TaskB = 9,
                  TaskC = 2,
                  TaskD = 0)
nPointQ1Total <- sum(unlist(lPointsQ1))
```
<!-- @ -->
<!-- <<PointsQ2, echo=FALSE>>= -->
```{r PointsQ2, echo=FALSE}
# Assign Points for Q2
lPointsQ2 <- list(TaskA = 8,
                  TaskB = 25,
                  TaskC = 17,
                  TaskD = 0)
nPointQ2Total <- sum(unlist(lPointsQ2))
```
<!-- @ -->
<!-- <<PointsQ3, echo=FALSE>>= -->
```{r PointsQ3, echo=FALSE}
# Assign Points for Q3
lPointsQ3 <- list(TaskA = 16,
                  TaskB = 6,
                  TaskC = 3,
                  TaskD = 0)
nPointQ3Total <- sum(unlist(lPointsQ3))
```
<!-- @ -->
<!-- <<PointsQ4, echo=FALSE>>= -->
```{r PointsQ4, echo=FALSE}
# Assign Points for Q4
lPointsQ4 <- list(TaskA = 2,
                  TaskB = 11,
                  TaskC = 8,
                  TaskD = 0)
nPointQ4Total <- sum(unlist(lPointsQ4))
```
```{r PointsQ5, echo=FALSE}
# Assign Points for Q5
lPointsQ5 <- list(TaskA = 0,
                  TaskB = 0,
                  TaskC = 0,
                  TaskD = 0)
nPointQ5Total <- sum(unlist(lPointsQ5))
```
<!-- @ -->
<!-- <<PointsTotal, echo=FALSE>>= -->
```{r PointsTotal, echo=FALSE}
nPointOverallTotal <- nPointQ1Total + nPointQ2Total + nPointQ3Total + nPointQ4Total + nPointQ5Total
```
<!-- @ -->


\thispagestyle{empty}

\fcolorbox{white}{white}{
	\centering \parbox[t]{1.0\linewidth}{
		\fontsize{12pt}{20pt}\selectfont % 
		\vspace*{0.5cm} % 

   	Peter von Rohr \\ Institute of Agricultural Sciences\\ D-USYS\\ ETH Zurich

		\vspace*{0.5cm} 
	}
}

\vspace*{2cm}

\fcolorbox{white}{white}{
	\parbox[t]{1.0\linewidth}{
		\centering \fontsize{25pt}{40pt}\selectfont %
		\vspace*{0.2cm}
		 751-6212-00L V \\
     Solutions to Exam    \\
     Applied Genetic Evaluation\\
     SS 2021

		\vspace*{0.7cm} % Space between the end of the title and the bottom of the grey box
	}
}


\vspace*{1cm}

<!-- % Table with Name -->
\begin{tabular}{p{3cm}p{6cm}}
Date:     &  $31^{st}$ May 2020 \\
          &  \\
          &  \\          
Name:     &  `r params$firstname` `r params$name` \\
          &  \\
          &  \\
Legi-Nr:  & `r params$leginr`\\
\end{tabular}

<!-- % Table with Points -->

\vspace{5ex}
\begin{center}
\begin{tabular}{|p{3cm}|c|c|}
\hline
Problem  &  Maximum Number of Points  &  Number of Points Reached\\
\hline
1        &  `r nPointQ1Total`  & \\
\hline
2        &  `r nPointQ2Total`  & \\
\hline
3        &  `r nPointQ3Total`  & \\
\hline
4        &  `r nPointQ4Total`  & \\
\hline
Total    &  `r nPointOverallTotal` & \\
\hline
\end{tabular}
\end{center}

\vspace{1cm}

\textit{Questions in German are in italics}

\clearpage
\pagebreak

## Problem 1: Breeding Program
A beef production herd reported the following economic results with respect to their beef fattening operation.

\textit{Ein Rindviehmastbetrieb weist die folgenden wirtschaftlichen Resultate aus dem Mastbereich aus.}

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
n_wean_weight <- 250
n_sla_weight <- 550
n_dress <- 0.6
n_agd <- 1
n_carcass_price <- 12.25
n_sub <- 2340
n_dir_cost <- 10 
n_fix_cost <- 2800
tbl_eco_result_p01 <- tibble::tibble(Parameter = c("Weaning Weight (_Absetzgewicht_)",
                                                   "Slaughter Weight (_Schlachtgewicht_)",
                                                   "Dressing Percentage (_Schlachtausbeute_)",
                                                   "Average Daily Gain (_Tageszunahme_)",
                                                   "Carcass Price (_Schlachtkörperpreis_)",
                                                   "Subsidies (_Subventionen_)",
                                                   "Direct Costs (_Direkte Kosten_)",
                                                   "Fixed Costs (_Fixe Kosten_)"),
                                     Value     = c(paste(n_wean_weight, "kg"),
                                                   paste(n_sla_weight, "kg"),
                                                   n_dress,
                                                   paste(n_agd,  "kg/day"),
                                                   paste(n_carcass_price, "SFr / [kg carcass weight]"),
                                                   paste(n_sub, "SFr / slaughter animal"),
                                                   paste(n_dir_cost, "SFr / slaughter animal and day"),
                                                   paste(n_fix_cost, "SFr / slaughter animal")))
knitr::kable(tbl_eco_result_p01)
```


\begin{enumerate}
\item[a)] Compute the profit per slaughter animal based on the economic parameters given in the table above.

\textit{Berechnen Sie den Gewinn pro Schlachttier aufgrund der oben gegebenen wirtschaftlichen Parameter.}
\end{enumerate}
\points{`r lPointsQ1$TaskA`}

### Solution

### Master Solution
The profit ($P$) per slaughter animal is given by the difference between revenue ($R$) and costs ($C$). The revenue corresponds to the sum of the price per slaughter animal plus the subsidices. The price per slaughter animal is the slaughter weight times the slaughter price.

The costs are the sum from fixed costs and direct costs. The direct costs depend on the length of the fattening period which is the difference in slaughter weight minus weaning weight divided by the average daily gain.

```{r, echo=TRUE, results='markup'}
# revenue
n_sl_price <- n_sla_weight * n_dress * n_carcass_price
n_rev <- n_sl_price + n_sub
# length of fattening period
n_len_fat_period <- (n_sla_weight - n_wean_weight) / n_agd
# direct costs
n_tot_dir_cost <- n_len_fat_period * n_dir_cost
# total costs
n_tot_cost <- n_tot_dir_cost + n_fix_cost
# profit
n_profit <- n_rev - n_tot_cost
```

The profit per slaughter animal is: `r n_profit` SFr.


\clearpage
\pagebreak


\begin{enumerate}
\item[b)] Compute the economic value for the traits average daily gain (AGD), fleshiness and fat coverage based on the values given in the table above. For the economic value of fleshiness and fat coverage, the price deductions per class as shown in the table below are assumed. 

\textit{Berechnen Sie die wirtschaftlichen Gewichte für Tageszunahme, Fleischigkeit und Fettabdeckung basierend auf den oben gemachten Angaben. Für die wirtschaftlichen Gewichte der Fleischigkeit und der Fettabdeckung werden die folgenden Preisabzüge pro Klasse angenommen.}
\end{enumerate}

\points{`r lPointsQ1$TaskB`}

```{r, echo=FALSE, results='asis'}
n_price_red_flesh <- 0.03
n_price_red_fat <- 0.02
tbl_price_red <- tibble::tibble(Parameter = c("Price Reduction for Fleshiness (_Preisabzug für Fleischigkeit_)",
                                              "Price Reduction for Fat Coverage (_Preisabzug für Fettabdeckung_)"),
                                Value     = c(paste(100*n_price_red_flesh, "percentage of carcass price"),
                                              paste(100*n_price_red_fat, "percentage of carcass price")))
```


### Solution 

### Master Solution 
```{r, echo=FALSE}
n_agd_incr <- 0.001
```

Economic values for the three traits are computed as follows

* ADG: Change of profit caused by an increase of ADG by `r n_agd_incr` kg/day

```{r, echo=TRUE, results='markup'}
n_agd_new <- n_agd + n_agd_incr
# length of fattening period
n_len_fat_period_new <- (n_sla_weight - n_wean_weight) / n_agd_new
# direct costs
n_tot_dir_cost_new <- n_len_fat_period_new * n_dir_cost
# total costs
n_tot_cost_new <- n_tot_dir_cost_new + n_fix_cost
# profit
n_profit_new <- n_rev - n_tot_cost_new
# economic value for agd
n_ev_agd <- n_profit_new - n_profit
```

* Fleshiness: The profit is only changed by the change in revue. The revenue is changed by the price deduction.

```{r}
# revenue
n_sl_price_new <- n_sla_weight * n_dress * n_carcass_price * (1-n_price_red_flesh)
n_rev_new <- n_sl_price_new + n_sub
# length of fattening period
n_len_fat_period <- (n_sla_weight - n_wean_weight) / n_agd
# direct costs
n_tot_dir_cost <- n_len_fat_period * n_dir_cost
# total costs
n_tot_cost <- n_tot_dir_cost + n_fix_cost
# profit
n_profit_new <- n_rev_new - n_tot_cost
# economic value for fleshiness
n_ev_fl <- n_profit - n_profit_new
```

* Fat coverage: the same as with fleshiness

```{r}
# revenue
n_sl_price_new <- n_sla_weight * n_dress * n_carcass_price * (1-n_price_red_fat)
n_rev_new <- n_sl_price_new + n_sub
# length of fattening period
n_len_fat_period <- (n_sla_weight - n_wean_weight) / n_agd
# direct costs
n_tot_dir_cost <- n_len_fat_period * n_dir_cost
# total costs
n_tot_cost <- n_tot_dir_cost + n_fix_cost
# profit
n_profit_new <- n_rev_new - n_tot_cost
# economic value for fleshiness
n_ev_fc <- n_profit - n_profit_new
```
* Result table

```{r, echo=FALSE, results='asis'}
vec_ev_result <- c(n_ev_agd, n_ev_fl, n_ev_fc)
tbl_ev_result <- tibble::tibble(Trait = c("ADG", "Fleshiness", "Fat Coverage"),
                                `Economic Value` = round(vec_ev_result, digits = 3))
knitr::kable(tbl_ev_result)
```


\clearpage
\pagebreak

\begin{enumerate}
\item[c)] For two bulls the following predicted breeding values are given. Which of the two bulls would you select based on the index value obtained from the aggregate genotype ($H$) containing the three traits: average daily gain, fleshiness and fatness? 

\textit{Für zwei Bullen sind die folgenden geschätzen Zuchtwerte gegeben. Welchen der beiden Bullen würden Sie aufgrund des Indexwertes des Gesamtzuchtwertes ($H$) mit den drei Mermalen Tageszunahme, Fleischigkeit und Fettabdeckung auswählen?}
\end{enumerate}

\points{`r lPointsQ1$TaskC`}

```{r, echo=FALSE, results='asis'}
vec_bv_bull1 <- c(250, -0.01, -0.02)
vec_bv_bull2 <- c(20, 0.02, 0.015)
tbl_pbv <- tibble::tibble(Trait = c("Average Daily Gain (g/day)",
                                    "Fleshiness (class)",
                                    "Fat Coverage (class)"),
                          `Bull 1` = vec_bv_bull1,
                          `Bull 2` = vec_bv_bull2)
knitr::kable(tbl_pbv)
```


### Solution 

### Master Solution 

* Bull 1: Index value

```{r, echo=TRUE, results='markup'}
(n_idx_val_bull1 <- crossprod(vec_bv_bull1, vec_ev_result))
```

* Bull 2: Index value

```{r, echo=TRUE, results='markup'}
(n_idx_val_bull2 <- crossprod(vec_bv_bull2, vec_ev_result))
```

```{r, echo=FALSE}
if (n_idx_val_bull1 > n_idx_val_bull2){
  s_decision <- "Bull 1 is better than Bull 2, hence we select Bull 1"
} else {
  s_decision <- "Bull 1 is not better than Bull 2, hence we select Bull s"
}
```

* Decision: `r s_decision`


\clearpage
\pagebreak

## Problem 2: Prediction of Breeding Values

```{r, echo=FALSE}
s_data_url_path <- "https://charlotte-ngs.github.io/gelasmss2021/data/gel_exam_p02.csv"
n_h2_imf <- 0.25
n_sigma_p2 <- 2.25
```

The following dataset will be used to predict breeding values using different methods. The dataset is available under the following URL:

\textit{Der nachfolgende Datensatz wird für die Schätzung von Zuchtwerten mit verschiedenen Methoden verwendet. Der Datensatz ist unter dem folgenden URL verfügbar:}

`r s_data_url_path`

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
tbl_data_p02 <- readr::read_csv2(file = s_data_url_path)
knitr::kable(tbl_data_p02)
```

### Hints
* Heritability (_Erblichkeit_) of the given trait is: `r n_h2_imf`
* Phenotypic variance (_Phänotypisch Varianz_): `r n_sigma_p2`

\vspace{3ex}
\begin{enumerate}
\item[a)] Predict the breeding values of the animals given in the table above using the given observation as an own-performance.

\textit{Schätzen Sie die Zuchtwerte der Tiere, welche in der Tabelle oben gegeben sind und verwenden Sie die angegebenen Beobachtungen als Eigenleistung.}
\end{enumerate}
\points{`r lPointsQ2$TaskA`}


### Solution

### Master Solution

The predicted breeding value ($\hat{u_i}$) for animal $i$ using own-performance is computed as

$$\hat{u_i} = h^2 (y_i - \mu)$$

The population mean $\mu$ is taken as the mean of all observations. 

```{r}
n_mu <- mean(tbl_data_p02$Observation)
n_hat_u <- n_h2_imf * (tbl_data_p02$Observation - n_mu)
tbl_result_pbv_p02a <- tibble::tibble(Animal = tbl_data_p02$Animal,
                                      Result = round(n_hat_u, digits = 3))
knitr::kable(tbl_result_pbv_p02a)
```


\clearpage
\pagebreak

\begin{enumerate}
\item[b)] Use the following dataset to predict breeding values using a BLUP animal model. Specify all model components, enter all information from the dataset into the model components and specify expected values and variance-covariance matrices for all random components in the model. 

\textit{Verwenden Sie den folgenden Datensatz zur Schätzung der Zuchtwerte mit einem BLUP-Tiermodell. Spezifizieren Sie alle Modellkomponenten, setzen Sie alle Informationen aus dem Datensatz in die Modellkomponenten ein und geben Sie die Erwartungswerte und die Varianz-Kovarianz-Matrizen der zufälligen Effekte im Modell an. }
\end{enumerate}
\points{`r lPointsQ2$TaskB`}


### Solution

### Master Solution
The BLUP animal model corresponds to the following mixed linear effect model

$$y = Xb + Zu + e$$

where $y$ is the vector of observations, $b$ is the vector of fixed effects, $u$ is the vector of random breeding values, $e$ is the vector of random residuals. The matrices $X$ and $Z$ are known incidence matrices which link the fixed effects and the random breeding values to the observations, respectively.

Inserting the data into the known components of the model leads to

```{r, echo=FALSE, results='asis'}
vec_y <- tbl_data_p02$Observation
n_nr_obs <- length(vec_y)
# design matrix X
tbl_data_p02$Sex <- as.factor(tbl_data_p02$Sex)
lm_fix <- lm(Observation ~ 0 + Sex, data = tbl_data_p02)
mat_X <- model.matrix(lm_fix)
attr(mat_X, which = "assign") <- NULL
attr(mat_X, which = "contrasts") <- NULL
# design matrix Z, founders in pedigree
vec_animals <- tbl_data_p02$Animal
vec_sire <- tbl_data_p02$Sire
vec_dam <- tbl_data_p02$Dam
vec_founder_sire <- setdiff(vec_sire, vec_animals)
vec_founder_dam <- setdiff(vec_dam, vec_animals)
vec_founder <- union(vec_founder_sire, vec_founder_dam)
vec_founder <- vec_founder[order(vec_founder)]
n_nr_founder <- length(vec_founder)
mat_Z <- cbind(matrix(data = 0, nrow = n_nr_obs, ncol = n_nr_founder),
               diag(1, nrow = n_nr_obs))

# unknowns
vec_b <- c("b_M", "b_F")
n_nr_ped_animal <- n_nr_founder + n_nr_obs
vec_u <- rmdhelp::vecGetVecElem(psBaseElement = "u", pnVecLen = n_nr_ped_animal, psResult = "latex")
vec_e <- rmdhelp::vecGetVecElem(psBaseElement = "e", pnVecLen = n_nr_obs, psResult = "latex")

# output
cat(paste0(rmdhelp::bcolumn_vector(pvec = vec_y, ps_name = "y", ps_env = "$$"), collapse = "\n"), "\n")
cat(paste0(rmdhelp::bmatrix(pmat = mat_X, ps_name = "X", ps_env = "$$"), collapse = "\n"), "\n")
cat(paste0(rmdhelp::bmatrix(pmat = mat_Z, ps_name = "Z", ps_env = "$$"), collapse = "\n"), "\n")

cat(paste0(rmdhelp::bcolumn_vector(pvec = vec_b, ps_name = "b", ps_env = "$$"), collapse = "\n"), "\n")
cat(paste0(rmdhelp::bcolumn_vector(pvec = vec_u, ps_name = "u", ps_env = "$$"), collapse = "\n"), "\n")
cat(paste0(rmdhelp::bcolumn_vector(pvec = vec_e, ps_name = "e", ps_env = "$$"), collapse = "\n"), "\n")
```

The expected values and the variance-covariance matrices for the random effects are

$$
E \left[\begin{array}{c}e \\ u \\ y \end{array} \right] = \left[\begin{array}{c}0 \\ 0 \\ Xb \end{array} \right]
$$

$$
var \left[\begin{array}{c}e \\ u \\ y \end{array} \right] 
= 
\left[\begin{array}{ccc} R & 0 & R\\ 0 & U & UZ^T\\ R & ZU & ZUZ^T + R\end{array} \right]
$$
with $R = I*\sigma_e^2$, $U = A * \sigma_u^2$ and $A$ corresponding to the numerator relationship matrix.

The mixed model equations are given by

$$
\left[ 
  \begin{array}{cc}
  X^TX  &  X^TZ  \\
  Z^TX  &  Z^TZ + A^{-1} * \lambda
  \end{array}
\right]
\left[ 
  \begin{array}{c}
  \hat{b} \\
  \hat{u}
  \end{array}
\right]
=
\left[ 
  \begin{array}{c}
  X^Ty  \\
  Z^Ty
  \end{array}
\right]
$$

The solutions to the mixed model equations are computed as follows

```{r, echo=TRUE, results='markup'}
lambda <- (1 - n_h2_imf) / n_h2_imf
# numerator relationship matrix
ped <- pedigreemm::pedigree(sire = c(rep(NA, n_nr_founder), tbl_data_p02$Sire), 
                            dam  = c(rep(NA, n_nr_founder), tbl_data_p02$Dam),
                            label = as.character(1:n_nr_ped_animal))
mat_Ainv <- as.matrix(pedigreemm::getAInv(ped = ped))
mat_XtX <- crossprod(mat_X)
mat_XtZ <- crossprod(mat_X, mat_Z)
mat_ZtX <- crossprod(mat_Z, mat_X)
mat_ZtZAinvlambda <- crossprod(mat_Z) + mat_Ainv * lambda
mat_coef <- rbind(cbind(mat_XtX, mat_XtZ), cbind(mat_ZtX, mat_ZtZAinvlambda))
mat_rhs <- rbind(crossprod(mat_X, vec_y), crossprod(mat_Z, vec_y))
mat_sol <- solve(mat_coef, mat_rhs)
```

The solution for the fixed effects are 

```{r, echo=FALSE, results='asis'}
# fixed effect solutons
n_nr_fix_effect <- ncol(mat_X)
vec_b_sol <- mat_sol[1:n_nr_fix_effect,]
# random effect
vec_u_sol <- mat_sol[(n_nr_fix_effect+1):nrow(mat_sol),]
# output
cat(paste0(rmdhelp::bcolumn_vector(pvec = round(vec_b_sol, digits = 4), ps_name = "\\hat{b}", ps_env = "$$"), collapse = "\n"), "\n")
cat(paste0(rmdhelp::bcolumn_vector(pvec = round(vec_u_sol, digits = 4), ps_name = "\\hat{u}", ps_env = "$$"), collapse = "\n"), "\n")
```



\clearpage
\pagebreak


\begin{enumerate}
\item[c)] Compare the Reliabilities of the predicted breeding values from a) and b). What is the reason for them being different?

\textit{Vergleichen Sie die Bestimmtheitsmasse der unter a) und b) geschätzten Zuchtwerte. Wo liegt der Grund für die unterschiedlichen Werte der Bestimmtheitsmasse?}
\end{enumerate}
\points{`r lPointsQ2$TaskC`}


### Solution 

### Master Solution 

Reliabilities ($B$) correspond to the squarred correlation ($r_{u, \hat{u}}^2$) between true and predicted breeding value. 

\clearpage
\pagebreak


## Problem 3: Model Selection

\textit{}

\begin{enumerate}
\item[a)] Model selection can be done using \textbf{forward} selection or \textbf{backward} selection. Describe the four steps that are needed in both model selection procedures.

\textit{Modellselektion kann mit \textbf{Forward} Selektion oder \textbf{Backward} Selektion gemacht werden. Beschreiben Sie die vier Schritte, welche es für beide Modellselektionsprozeduren braucht.}
\end{enumerate}
\points{`r lPointsQ3$TaskA`}


### Solution

\begin{tabular}{|l|p{6cm}|p{6cm}|}
\hline
  &  &  \\
Step  &  Forward   &  Backward  \\
  &  &  \\
\hline
  &  &  \\
1  &  
   &    \\
   &  &  \\
\hline
   &  &  \\
2  &  
   &   \\
   &  &  \\
\hline
   &  &  \\
3  &  
   &   \\
   &  &  \\
\hline
   &  &  \\
4  &  
   &   \\
   &  &  \\
\hline
\end{tabular}



### Master Solution

\begin{tabular}{|l|p{6cm}|p{6cm}|}
\hline
  &  &  \\
Step  &  Forward   &  Backward  \\
  &  &  \\
\hline
  &  &  \\
1  &  Start with smallest model $\mathcal{M}_0$ 
   &  Start with full model $\mathcal{M}_0$ \\
   &  &  \\
\hline
   &  &  \\
2  &  Include predictor reducing RSS the most
   &  Eliminate predictor increasing RSS the least \\
   &  &  \\
\hline
   &  &  \\
3  &  Continue step 2 until all predictors chosen
   &  Continue step 2 until all predictors eliminated  \\
   &  &  \\
\hline
   &  &  \\
4  &  From sequence of submodels $\mathcal{M}_0 \subseteq \mathcal{M}_1 \subseteq \mathcal{M}_2 \subseteq \ldots$ choose model with smallest $C_p$
   &  From sequence of submodels $\mathcal{M}_0 \supseteq \mathcal{M}_1 \supseteq \mathcal{M}_2 \supseteq \ldots$ choose model with smallest $C_p$ \\
   &  &  \\
\hline
\end{tabular}


\clearpage
\pagebreak


\begin{enumerate}
\item[b)] We are given the following data set on body weight, breast circumference and shoulder height. Furthermore, the results of fitting the full regression model and a reduced model with only breast circumference as predictor. Compute the Mallow $C_p$ value for both models and decide which of the two models is better based on $C_p$

\textit{Gegeben ist der folgende Datensatz zu Körpergewicht, Brustumfang und Widerristhöhe. Weiter sind auch die Resultate des Fits des vollen Modells und eines reduzierten Models mit nur Brustumfang als beschreibende Variable gegeben. Berechnen Sie den Mallow $C_p$-Wert für beide Modelle und begründen Sie, welches der beiden Modelle besser ist.}
\end{enumerate}

\points{`r lPointsQ3$TaskB`}

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
s_data_root_online <- "https://charlotte-ngs.github.io/gelasmss2021"
s_data_file <- "gel_exam_p03.csv"
s_data_p03_url_path <- file.path(s_data_root_online, "data", s_data_file)
tbl_data_p03 <- readr::read_csv2(file = s_data_p03_url_path)
knitr::kable(tbl_data_p03)
```

The dataset is available under:

\textit{Der Datensatz ist verfügbar unter: }

`r s_data_p03_url_path`

The results of the full model

\textit{Die Resultate des vollen Modells}

```{r FullModelP01b, echo=FALSE, results='markup'}
lm_bw_bc_sh <- lm(`BodyWeight` ~ `BreastCircumference` + `ShoulderHeight`, data = tbl_data_p03)
(sy_bw_bc_sh <- summary(lm_bw_bc_sh))
```

\clearpage
\pagebreak

The results of the reduced model

\textit{Die Resultate des reduzierten Modells}

```{r RedModelP01b, echo=FALSE, results='markup'}
lm_bw_bc <- lm(`BodyWeight` ~ `BreastCircumference`, data = tbl_data_p03)
(sy_bw_bc <- summary(lm_bw_bc))
```


### Solution

### Master Solution

Mit `olsrr` erhalten wir

\tiny

```{r SolP01b, echo=TRUE, results='markup'}
olsrr::ols_step_best_subset(lm_bw_bc_sh)
```

\normalsize

Das reduzierte Model erhält einen tieferen $C_p$-Wert und ist somit das bessere Modell.


\clearpage
\pagebreak

\begin{enumerate}
\item[c)] What is the reason why we are doing model selection, why should we not take always the full model? Please explain the underlying phenomenon in your own words.

\textit{Was ist der Grund, dass wir Modellselektion machen? Weshalb können wir nicht einfach immer das volle Modell mit allen erklärenden Variablen zur Anpassung verwenden? Bitte erklären Sie das unterliegende Phänomen in Ihren eigenen Worten.}
\end{enumerate}
\points{`r lPointsQ3$TaskC`}


### Solution

### Master Solution
The problem that makes model selection necessary is called Bias-Variance trade-off. The Bias is reduced, the more predictors we include. But on the other hand, the variance is increased. Hence we have to find the model that balances this trade-off. This is done with a criterion such as the Mallow $C_p$ statistic.

\clearpage
\pagebreak


## Problem 4: Variance Components Estimation

The following data set with average amounts of methane gas emission per year for three offspring per bull are given in the following table.

\textit{Im folgenden Datensatz sind die mittleren Mengen an Methanemissionen pro Jahr für je drei Nachkommen für fünf verschiedene Bullen gegeben.}

```{r TableDataP04, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
s_data_p04_file <- "gel_exam_p04.csv"
s_data_root_online <- "https://charlotte-ngs.github.io/gelasmss2021"
s_data_p04_path <- file.path(s_data_root_online, "data", s_data_p04_file)
tbl_met <- readr::read_csv2(file = s_data_p04_path)
knitr::kable(tbl_met,
             booktabs = TRUE,
             longtable = TRUE)
```

The data set is available under:

\textit{Der Datensatz ist verfügbar unter:}

`r s_data_p04_path`


\begin{enumerate}
\item[a)] Given that we want to reduce the amount of methane gas emitted in cattle using the tools of livestock breeding, we first have to do a variance components estimation. Why is it important to have a certain variability in a given trait and which variance component is important when we want to improve a certain trait with livestock breeding tools?

\textit{Wir möchten die Methanemission beim Rind mit den Werkzeugen der Tierzucht senken. Weshalb brauchen wir für die züchterische Bearbeitung eines Merkmals ein gewisse Variabilität im Merkmal und welche Varianzkomponente interessiert uns besonders?}
\end{enumerate}
\points{`r lPointsQ4$TaskA`}


### Solution

### Master Solution
One of the main livestock breeding tools is selection. Selection means that we choose from a given population the best individuals to be parents of the next generation. Hence selection can only happen, when a certain level of variation can be observed in a population. Because livestock breeding means improvement of a population at the genetic level, the genetic variance component or in our case the sire variance component is important.

\clearpage
\pagebreak


\begin{enumerate}
\item[b)] What is the linear model that would be used to allows us a separation of the different variance components in the given dataset on methane gas emission?

\textit{Wie sieht das lineare Modell aus, welches eine Aufteilung der Varianz in ihre verschiedenen Komponenten im Methandatensatz erlaubt?}
\end{enumerate}
\points{`r lPointsQ4$TaskB`}


### Solution

### Master Solution
The model is

$$y_{ij} = \mu + s_i + e_{ij}$$

\begin{tabular}{lll}
where  &  &  \\
       & $y_{ij}$         &  measurement $j$ of animal $i$    \\
       & $\mu$            &  expected value of $y$            \\
       & $s_i$            &  deviation of $y_{ij}$ from $\mu$ attributed to bull $i$ \\
       & $e_{ij}$  &  measurement error
\end{tabular}

The expected values are: $E\left[ s_i \right] = 0$, $E\left[ e_{ij} \right] = 0$, $E\left[ y_{ij} \right] = \mu$

The variances: $var(s_i) = \sigma_s^2$, $var(e_{ij}) = \sigma_e^2$, $var(y_{ij}) = \sigma_s^2 + \sigma_e^2$

\clearpage
\pagebreak

\begin{enumerate}
\item[c)] Use an analysis of variance (ANOVA) to estimate the variance components given in the above data set.

\textit{Schätzen Sie mit einer Varianzanalyse (ANOVA) die Varianzkomponenten aus dem oben angegebenen Datensatz zu den Methanemissionen.}
\end{enumerate}
\points{`r lPointsQ4$TaskC`}


### Solution

### Master Solution
The sires are first converted into factors to prevent R to fit them as covariables.

```{r PrepP03c, echo=TRUE, results='markup'}
tbl_met$Bull <- as.factor(tbl_met$Bull)
```

Use `aov()` to get the ANOVA-table

```{r AovP03c, echo=TRUE, results='markup'}
aov_meth <- aov(Methane ~ Bull, data = tbl_met)
(sy_aov_meth <- summary(aov_meth))
```

From this the estimate of the residual variance corresponds to

```{r Sigmae2, echo=TRUE, results='markup'}
(n_hat_sigma_e2 <- sy_aov_meth[[1]]$`Mean Sq`[2])
```

```{r Sigmas2, echo=TRUE, results='markup'}
n_nr_dau <- 3
(n_hat_sigma_s2 <- (sy_aov_meth[[1]]$`Mean Sq`[1] - sy_aov_meth[[1]]$`Mean Sq`[2]) / n_nr_dau)
```

