---
title: "Prepare Docker Container"
author: "Peter von Rohr"
date: "2/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Disclaimer
The preparation of the docker container for the R exercise platform is documented here.


## Dockerfile
The docker container that is built is based on the container `rocker/verse`. To have all packages available, we build a `Dockerfile` that installs all packages directly into the image. The content of the `Dockerfile` is shown below

```
$ cd /home/quagadmin/courses/gelasmss2021/docker
quagadmin@2-htz:~/courses/gelasmss2021/docker$ cat Dockerfile 
FROM rocker/verse

LABEL maintainer="peter.vonrohr@usys.ethz.ch"

RUN R -e "install.packages(c('pedigreemm', 'kableExtra', 'AlphaSimR', dependencies = TRUE))" && \
    R -e "remotes::install_github(repo = 'charlotte-ngs/rmdhelp')" && \
    R -e "remotes::install_github(repo = 'pvrqualitasag/rvcetools')"

```

From that `Dockerfile`, a docker image is built by the following command

```
docker build -t gelasmss2021 .
```


## Testing The Image
The following command is used to test the command

```
docker run -d -p 10087:8787 -e PASSWORD=pvrpass -v /home/quagadmin/courses/gelasmss2021/home/pvr:/home/rstudio --name pvr_rstudio gelasmss2021
$ docker ps -a
CONTAINER ID        IMAGE               COMMAND             CREATED             STATUS              PORTS                     NAMES
8e3d27e3799c        gelasmss2021        "/init"             9 seconds ago       Up 6 seconds        0.0.0.0:10087->8787/tcp   pvr_rstudio
```

Open the firewall to test it from outside

```
sudo ufw allow 10087/tcp
sudo ufw status
```

## Stopping the Test Container

```
docker stop pvr_rstudio
docker rm $(docker ps --filter "status=exited" -q)
```


## Creating Student Lists
Start with a download of the `BelegungLerneinheit*.xlsx` file from `eDoz` under `https://www.lehrbetrieb.ethz.ch/edoz` . We assume that this file is saved to the `Downloads` folder. From there, we extract the usernamens of the students.

```{r}
vec_pkg <- c('openxlsx', 'dplyr', 'tidyr')
if (!all(is.element(vec_pkg, installed.packages())))
  install.packages(vec_pkg, dependencies = TRUE)

s_bl_path <- '~/Downloads/BelegungenLerneinheit_751760200L_Fruehjahrssemester_2021.xlsx'
wb <- openxlsx::read.xlsx(xlsxFile = s_bl_path)
```

First create a list of students with their names, firstnames and e-mail addresses

```{r}
require(dplyr)
(tbl_student_info <- wb %>% 
    select(Familienname, Rufname, Nummer, `E-Mail`) %>%
    mutate(UDName = `E-Mail`) %>%
    tidyr::separate(UDName, c("Username", NA), sep = '@') %>%
    select(Username, Familienname, Rufname, Nummer, `E-Mail`))
```

The student information is written to a file. 

```{r}
readr::write_csv(tbl_student_info, file = file.path(here::here(), 'students', 'students_gelasmss2021.txt'), col_names = FALSE)
```

A test-student is created manually. 

```{r}
tbl_test_student_info <- tibble::tibble(Username     = c('vrohrp'),
                                        Familienname = c('von Rohr'), 
                                        Rufname      = c('Peter'), 
                                        Nummer       = c('00-000-000') , 
                                        `E-Mail`     = c('peter.vonrohr@usys.ethz.ch'))
readr::write_csv(tbl_test_student_info, file = file.path(here::here(), 'students', 'test_students_gelasmss2021.txt'), col_names = FALSE)
```



## Student Usernames
From the e-mail addresses, we extract usernames as follows

```{r}
(tbl_student_username <- tbl_student_info %>% select(Username))
```

Write the usernames to a file

```{r}
readr::write_csv(tbl_student_username, file = file.path(here::here(), 'students', 'student_usernames_gelasmss2021.txt'), col_names = FALSE)
```

The username for the test-student is specified manually and is not extracted from the e-mail. 

```{r}
tbl_test_student_username <- tbl_test_student_info %>% select(Username)
readr::write_csv(tbl_test_student_username, file = file.path(here::here(), 'students', 'test_student_usernames_gelasmss2021.txt'), col_names = FALSE)
```


## File Upload
Upload the created files to the server

```
$ cd /Users/peter/Data/Projects/GitHub/charlotte-ngs/gelasmss2021_gh-root/master/gelasmss2021
$ scp -r students quagadmin@2-htz.quagzws.com:/home/quagadmin/courses/gelasmss2021
$ scp -r bash quagadmin@2-htz.quagzws.com:/home/quagadmin/courses/gelasmss2021
```


## Start Containers

```
cd /home/quagadmin/courses/gelasmss2021/bash
# test
./docker_student_start.sh -f ../students/test_students_gelasmss2021.txt -c /home/quagadmin/courses/gelasmss2021/home -p 10187
# real
./docker_student_start.sh -f ../students/students_gelasmss2021.txt -c /home/quagadmin/courses/gelasmss2021/home
```


## Cloning Material
The exercise material must be cloned to the directories of the students. First start with preparing the branch.

```
# on mac
cd /Users/pvr/Data/Projects/Github/charlotte-ngs/gelasmss2021_gh-root
mkdir -p rexpf
cd rexpf
git clone https://github.com/charlotte-ngs/gelasmss2021.git
cd rexpf
git branch rexpf
git checkout rexpf
ls -1 | while read d
do
  echo " * Checking element: $d ..."
  if [ "$d" != "ex" ] 
  then
    echo " ** Deleting $d ..."
    rm -rf $d
  fi
  sleep 2
done
545  git add .
546  git commit -m"Deleted all unneeded material"
547  git push origin rexpf
```

Cloning the material

```
# on 2-htz
export GITURI=https://github.com/charlotte-ngs/gelasmss2021.git
export GELSRC=/home/quagadmin/courses/gelasmss2021
export GELHOME=${GELSRC}/home
cd $GELSRC
# test student
cat students/test_student_usernames_gelasmss2021.txt | while read s
do
  cd $GELHOME/$s
  echo " * Cloning repo for $s ..."
  git clone $GITURI -b rexpf
  sleep 2
done

# students
cd $GELSRC
# test student
cat students/student_usernames_gelasmss2021.txt | while read s
do
  cd $GELHOME/$s
  echo " * Cloning repo for $s ..."
  git clone $GITURI -b rexpf
  sleep 2
done
```

## Sending E-Mails
The following commands are used to send e-mails to students

```
export GELSRC=/home/quagadmin/courses/gelasmss2021
cd $GELSRC
if [ ! -d "email_sent" ]; then mkdir -p email_sent;fi
cat students/test_students_gelasmss2021.txt  | while read s
do
  user=$(echo "$s" | cut -d ',' -f1)
  email=$(echo "$s" | cut -d ',' -f5)
  echo " * Sending ${user}_mail.txt to $email ..."
  cat bash/${user}_email.txt | ssmtp $email
  sleep 2
  mv bash/${user}_email.txt email_sent
done

# real students
cat students/students_gelasmss2021.txt  | while read s
do
  user=$(echo "$s" | cut -d ',' -f1)
  email=$(echo "$s" | cut -d ',' -f5)
  echo " * Sending ${user}_mail.txt to $email ..."
  cat bash/${user}_email.txt | ssmtp $email
  sleep 2
  mv bash/${user}_email.txt email_sent
done

```


# Restart Containers
For an update of the docker image, we have to restart the docker container. First the running container are stopped. 

```
# test
cd /home/quagadmin/courses/gelasmss2021/bash
# stop
./docker_student_stop.sh -f ../students/test_students_gelasmss2021.txt
# restart
./docker_student_start.sh -f ../students/test_students_gelasmss2021.txt -c /home/quagadmin/courses/gelasmss2021/home -p 10187 -w /home/quagadmin/courses/gelasmss2021/created -i gelasmss2021
```

The same can be done for all other students

```
# real
cd /home/quagadmin/courses/gelasmss2021/bash
# stop
./docker_student_stop.sh -f ../students/students_gelasmss2021.txt
# restart
./docker_student_start.sh -f ../students/students_gelasmss2021.txt -c /home/quagadmin/courses/gelasmss2021/home -w /home/quagadmin/courses/gelasmss2021/created -i gelasmss2021
```



